Lesson 5
========================================================

### Multivariate Data
Notes: Comparing data using 3 or more variables

***

### Moira Perceived Audience Size Colored by Age
Notes: Methods of adding more degrees: Color?

Plot may not reveal anything at all.

***

### Third Qualitative Variable
Notes:

```{r Third Qualitative Variable}
library('ggplot2')
pf <- read.csv('pseudo_facebook.tsv', sep='\t')
ggplot(aes(x = gender, y = age),
       data = subset(pf, !is.na(gender))) + geom_boxplot() +
  stat_summary(fun.y=mean, geom='point', shape=4)

ggplot(aes(x = age, y = friend_count),
       data = subset(pf, !is.na(gender))) +
  geom_line(aes(color=gender), stat="summary", fun.y=median)
```

***

### Plotting Conditional Summaries
Notes:

```{r Plotting Conditional Summaries}
library("dplyr")

pf.fc_by_age_gender <- pf %.%
  group_by(age, gender) %.%
  summarize(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %.%
  arrange(age, gender)

pf.fc_by_age_gender <- subset(pf, !is.na(gender)) %.%
  group_by(age, gender) %.%
  summarize(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %.%
  arrange(age, gender)

pf.fc_by_age_gender <- ungroup(pf.fc_by_age_gender)

# Instructor Method:
pf.fc_by_age_gender <- pf %.%
  filter(!is.na(gender)) %.%
  group_by(age, gender) %.%
  summarize(mean_friend_count = mean(friend_count),
            median_friend_count = median(friend_count),
            n = n()) %.%
  ungroup() %.%
  arrange(age)

ggplot(aes(y=median_friend_count, x=age), data=pf.fc_by_age_gender) +
  geom_line(aes(color=gender))
```

***

### Thinking in Ratios
Notes:

***

### Wide and Long Format
Notes: 

***

### Reshaping Data
Notes:

```{r}
# install.packages('reshape2')
library(reshape2)

pf.fc_by_age_gender.wide <- dcast(pf.fc_by_age_gender,
                                 age ~ gender,
                                 value.var  = "median_friend_count")

ggplot(data=pf.fc_by_age_gender.wide) +
  geom_line(aes(x=age, y=(female/male), color="pink")) +
  geom_hline(yintercept=1, color="black", linetype=2, alpha=.3)
```


***

### Ratio Plot
Notes:

```{r Ratio Plot}

```

***

### Third Quantitative Variable
Notes:

```{r Third Quantitative Variable}
pf$year_joined <- 2014 - ceiling((pf$tenure/365))
```

***

### Cut a Variable
Notes:

```{r Cut a Variable}
pf$year_joined.bucket <- cut(pf$year_joined, breaks=c(2004,2009,2011,2012,2014)) # (2004,2009],(2010,2011] ...
```

***

### Plotting it All Together
Notes:

```{r Plotting it All Together}
ggplot(aes(x=age, y=friend_count), data=subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=median )
```

***

### Plot the Grand Mean
Notes:

```{r Plot the Grand Mean}
ggplot(aes(x=age, y=friend_count), data=subset(pf, !is.na(year_joined.bucket))) + 
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=mean ) + 
  geom_line(linetype=2, stat='summary', fun.y=mean)

```

***

### Friending Rate
Notes:

```{r Friending Rate}
pf$friend_rate <- (pf$friend_count / pf$tenure)
with(subset(pf, tenure > 0), summary(friend_count / tenure))

```

***

### Friendships Initiated
Notes:

What is the median friend rate?
.22
What is the maximum friend rate?
417.0
```{r Friendships Initiated}
ggplot(aes(x=tenure, y=(friendships_initiated / tenure)), data=subset(pf, tenure>0)) + 
  geom_line(aes(color=year_joined.bucket), stat='summary', fun.y=mean)
```

***

### Bias-Variance Tradeoff Revisited
Notes:

```{r Bias-Variance Tradeoff Revisited}

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure >= 1)) +
  geom_line(aes(color = year_joined.bucket),
            stat = 'summary',
            fun.y = mean)

ggplot(aes(x = 7 * round(tenure / 7), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 30 * round(tenure / 30), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = 90 * round(tenure / 90), y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_line(aes(color = year_joined.bucket),
            stat = "summary",
            fun.y = mean)

ggplot(aes(x = tenure, y = friendships_initiated / tenure),
       data = subset(pf, tenure > 0)) +
  geom_smooth(aes(color = year_joined.bucket))

```

***

### Sean's NFL Fan Sentiment Study
Notes:

***

### Introducing the Yogurt Data Set
Notes:

***

### Histograms Revisited
Notes:

```{r Histograms Revisited}
yo <- read.csv("yogurt.csv")
str(yo)

ggplot(aes(price), data=yo) +
  geom_histogram(binwidth=10)

summary(yo)
unique(yo$price)
length(unique(yo$price))
table(yo$price)
```

***

### Number of Purchases
Notes:

```{r Number of Purchases}

yo <- transform(yo, all.purchases = strawberry + blueberry + pina.colada + plain + mixed.berry)
```

***

### Prices over Time
Notes:

```{r Prices over Time}
ggplot(aes(x=time, y=price), data=yo) +
  geom_point(alpha=.05)
```

***

### Sampling Observations
Notes:

***

### Looking at Samples of Households

```{r Looking at Sample of Households}
library(dplyr)
set.seed(seed=8896)
sample.ids <- sample(levels(yo$id), 16)

ggplot(aes(x=time, y=price), 
       data=subset(yo, id %in% sample.ids)) + 
  facet_wrap(~ id) + 
  geom_line() + 
  geom_point(aes(size = all.purchases), pch=1)

ggsave('facetHouse.png')


```
In my plot subset I seem to have about a third of the graphs have examples of decreasing price through time, but the overall tendency is is for most of the points to be clustered around 60 for the whole duration. On the other hand, those with prices focused around 60, they almost all show examples of sharp price decreases for specific dates, seeming to indicated as mentioned, maybe specific sales.
***

### The Limits of Cross Sectional Data
Notes: Good for instances of categorical variables over time

***

### Many Variables
Notes: Might be able to iterate over different variable correspondences to discover relationship where none were thought to exist before

***

### Scatterplot Matrix
Notes:
Iterates over all pairs of variables to potentially unearth relationships with graphs

```{r}
install.packages('GGally')
library(GGally)
theme_set(theme_minimal(20))
pf<-read.csv('pseudo_facebook.tsv', sep='\t')
set_seed(1836)
pf_subset <- pf[ , c(2:7)]
names(pf_subset)
# ggpairs(pf_subset[sample.int(nrow(pf_subset), 1000), ])
ggpairs(pf_subset[sample.int(99003, 1000), ])
```
***

### Even More Variables
Notes:

***

### Heat Maps
Notes:

```{r}
nci <- read.table("nci.tsv")
colnames(nci) <- c(1:64)
```

```{r}
library('dplyr')
library('reshape2')
nci.long.samp <- melt(as.matrix(nci[1:200,]))
names(nci.long.samp) <- c("gene", "case", "value")
head(nci.long.samp)

ggplot(aes(y = gene, x = case, fill = value),
  data = nci.long.samp) +
  geom_tile() +
  scale_fill_gradientn(colours = colorRampPalette(c("blue", "red"))(100))
```


***

### Analyzing Three of More Variables
Reflection:

***

Click **KnitHTML** to see all of your hard work and to have an html
page of this lesson, your answers, and your notes!

